### 2019-06-18 2-Chapters/1-Chapter/chapter1.tex
Lots of details on Part I of the thsis

We present in Section~\ref{sec:4:firstModel} a first model of an IoT network, that is made of up-to thousands of end-devices and a single base station.
We consider a time-frequency slotted ALOHA-based protocol, by assuming a perfect synchronization between all devices and the base station.
%
Our IoT network model contains two types of devices: \emph{static devices} that use only one channel (fixed in time), and \emph{dynamic devices} that can choose the channel for each of their transmissions.
Static devices form an interfering traffic, which could be generated by devices using other standards as well.
Dynamic devices can use RL, and especially MAB algorithms, to optimize their sequential choices of channels, in order to maximize the number of successful up-link transmissions, thanks to the success/failure feedback that is the presence/absence of acknowledgement (\Ack) from the base station.
%
We first detail three base-lines, a naive approach and two centralized ``oracle'' approaches,
that are used to evaluate the performance of MAB algorithms (\UCB{} and TS), in terms of the centralized system-wide successful communication rate.
We show that our approach is interesting, as these low-cost MAB algorithms have near-optimal performance, even when the proportion of end-devices increases and the interfering traffic from other devices becomes more and more unpredictable.
%
Even without changing anything of the level of IoT standards, our proposal is just an add-on capability that can be used on a unit-per-unit basis.
This model constitutes the first article written during this thesis, \cite{Bonnefoi17}, with R. Bonnefoi, another PhD student of our team SCEE.

Verifying the potential gain of performance brought by these MAB algorithms in numerical simulations is a first step, but it is satisfactory to also verify our results on a real-world proof-of-concept (PoC).
In Section~\ref{sec:4:firstModel}, we propose the first PoC of using decentralized MAB learning in a simplified ``IoT'' like wireless network \cite{Besson2018ICT}.
We were able to implement a network with a base station which receives incoming messages from different end-devices and replies with acknowledgements, that are used as feedback by the devices and let them learn an efficient spectrum access scheme.
% Verify this first model empirically: proof of concept

We then consider extension of our first model, following the idea of packet retransmission that is at the root of the ALOHA protocol \cite{Abramson1970,Roberts75}.
This second model proposes to also use decision making at the second layer of the MAC protocol, that is to also use RL and MAB algorithms, in order to learn the best channel to use for retransmitting any packet that suffered from a collision at its first transmission.
We present the model in Section~\ref{sec:4:retransmissions}, along with mathematical justifications of the potential of using this second-stage MAB learning for retransmissions, and numerical illustrations \cite{Bonnefoi2019WCNC}.
This shows that using learning in a decentralized way on the devices' side is again a powerful idea under this model with retransmissions, and we show as well as the simplest of the different heuristics we compared is efficient and allow the devices to learn to behave almost optimally.

\TODOL{Je dois faire ça d'ici mercredi matin}

\paragraph{Chapter~\ref{chapter:5}.}
%
In the two models presented before, the core concept is to let all the dynamic devices of a IoT network run a fully decentralized learning algorithm,
each of them targeting their local objective of maximizing their cumulated reward (in a selfish way),
in order to optimize the complete system.
It is well known in the game theory literature that playing selfishly can be disastrous for the centralized performance measure: for instance, one can think of the prisoner dilemma.
Selfish
\TODOL{Je dois faire ça d'ici mercredi matin}

@@

\paragraph{Our approach.}
%
Our approach for the aforementioned problems is summarized as follows:
we propose a couple of new models of decision making for IoT networks, on the device's side, for the spectrum access problem.
By considering the difference between the hypothesis of previously studied models for the OSA problem and our problem at end for IoT, we identify the constraints and specificities of applying decentralized ML for IoT devices.
%
Like for the OSA, we can apply successfully model the decision making problem as a Multi-Armed Bandit problem.
We showed on numerical simulations as well as on a real-world proof-of-concept that applying low-cost sequential learning algorithms coming from the MAB literature successfully reduce the collision rate of IoT end-devices.
We proposed new algorithms, some of which are proven to converge efficiently.

@@

% WARNING
\newpage

>>>>

We present in Section~\ref{sec:4:firstModel} a first model of an IoT network, that is made of up-to thousands of end-devices and a single base station.
We consider a time-frequency slotted ALOHA-based protocol, by assuming a perfect synchronization between all devices and the base station.
%
Our IoT network model contains two types of devices: \emph{static devices} that use only one channel (fixed in time), and \emph{dynamic devices} that can choose the channel for each of their transmissions.
Static devices form an interfering traffic, which could be generated by devices using other standards as well.
Dynamic devices can use RL, and especially MAB algorithms, to optimize their sequential choices of channels, in order to maximize the number of successful up-link transmissions, thanks to the success/failure feedback that is the presence/absence of acknowledgement (\Ack) from the base station.
%
We first detail three base-lines, a naive approach and two centralized ``oracle'' approaches,
that are used to evaluate the performance of MAB algorithms (\UCB{} and TS), in terms of the centralized system-wide successful communication rate.
We show that our approach is interesting, as these low-cost MAB algorithms have near-optimal performance, even when the proportion of end-devices increases and the interfering traffic from other devices becomes more and more unpredictable.
%
Even without changing anything of the level of IoT standards, our proposal is just an add-on capability that can be used on a unit-per-unit basis.
This model constitutes the first article written during this thesis with R. Bonnefoi, another PhD student of our team SCEE.
We presented this first work, which obtained the ``best paper award'', at the EAI CROWNCOM conference in Lisboa, Portugal in September $2017$ \cite{Bonnefoi17}.

Verifying the potential gain of performance brought by these MAB algorithms in numerical simulations is a first step, but it is satisfactory to also verify our results on a real-world proof-of-concept (PoC).
In Section~\ref{sec:4:firstModel}, we propose the first PoC of using decentralized MAB learning in a simplified ``IoT'' like wireless network.
We were able to implement a network with a base station which receives incoming messages from different end-devices and replies with acknowledgements, that are used as feedback by the devices and let them learn an efficient spectrum access scheme.
% Verify this first model empirically: proof of concept
This work was again implemented with R. Bonnefoi, and was demonstrated during three days at the IEEE ICT conference in Saint-Malo, France, in June $2018$ \cite{Besson2018ICT}.

We then consider an extension of our first model, following the idea of packet retransmission that is at the root of the ALOHA protocol \cite{Abramson1970,Roberts75}.
This second model proposes to also use decision making at the second layer of the MAC protocol, that is to also use RL and MAB algorithms, in order to learn the best channel to use for retransmitting any packet that suffered from a collision at its first transmission.
We present the model in Section~\ref{sec:4:retransmissions}, along with mathematical justifications of the potential of using this second-stage MAB learning for retransmissions, and numerical illustrations.
This shows that using learning in a decentralized way on the devices' side is again a powerful idea under this model with retransmissions, and we show as well as the simplest of the different heuristics we compared is efficient and allow the devices to learn to behave almost optimally.
This third contribution concluded our collaboration with R. Bonnefoi, and was presented at the MoTION workshop during the IEEE WCNC conference in Marrakech, Morocco, in April $2019$ \cite{Bonnefoi2019WCNC}.

\paragraph{Chapter~\ref{chapter:5}.}
%
In the two models presented above, the core concept is to let all the dynamic devices of a IoT network run a fully decentralized learning algorithm,
each of them targeting their local objective of maximizing their cumulated reward (in a selfish way), in order to optimize the complete system.
It is well known in the game theory literature that playing selfishly can be disastrous for the centralized performance measure: for instance, one can think of the prisoner dilemma.
So it is quite surprising that our numerical simulations as well as the realistic PoC showed that decentralized selfish MAB learning lead to efficient

Selfish
\TODOL{Je dois faire ça d'ici mercredi matin}

@@

% \paragraph{Our approach.}
% %
% Our approach for the aforementioned problems is summarized as follows:
% we propose a couple of new models of decision making for IoT networks, on the device's side, for the spectrum access problem.
% By considering the difference between the hypothesis of previously studied models for the OSA problem and our problem at end for IoT, we identify the constraints and specificities of applying decentralized ML for IoT devices.
% %
% Like for the OSA, we can apply successfully model the decision making problem as a Multi-Armed Bandit problem.
% We showed on numerical simulations as well as on a real-world proof-of-concept that applying low-cost sequential learning algorithms coming from the MAB literature successfully reduce the collision rate of IoT end-devices.
% We proposed new algorithms, some of which are proven to converge efficiently.

@@

% WARNING
% \newpage
